$K$-corrections and filter transformations \\ in the ultraviolet, optical, and near infrared

### 2006-08-17 docs/paper/kcorrect-aj-v2/ms.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@563 75dff457-8751-0410-ab1a-cd701654ce6b

We begin with a basis set of 485 spectral templates. Of these, 450 are
a set of instantaneous bursts from \citet{bruzual03a}, using the
\citet{chabrier03a} stellar initial mass function and the Padova1994
isochrones. We select 6 metallicities (0.005, 0.02, 0.2, 0.4, 1, and
2.5 times solar). For each metallicity we select 25 ages (between 1
Myr and 13.75 Gyr, spaced almost logarithmically in age).  For each
age and metallicity we make 3 choices of dust model: (1) no dust
extinction; (2) $\tau_V = 3$ dust with Milky Way type extinction (as
input into the models of
\citealt{witt00a} with a ``homogeneous'' distribution and ``shell''
geometry); (3) $\tau_V = 3$ dust with SMC type extinction (from the
same models). We smooth each template to $300$ km s$^{-1}$ velocity
dispersion (we will smooth the SDSS spectra in the training set to the
same resolution).

The remaining 35 templates are from MAPPINGS-III (\citealt{kewley01a})
models of emission from ionized gas. We choose the predictions for an
8 Myr old continuous star-formation history with 5 possible
metalliciticites (0.5, 0.2, 0.4, 1.0, and 2. times solar) and 7
possible ionization parameters ($q= 5\times 10^6$, $10^7$, $2\times
10^7$, $4\times 10^7$, $8 \times 10^7$, $1.5\times 10^8$, and $3\times
10^8$ cm s$^{-1}$). We take the spectra (given as a set of delta
functions) and smooth them to $300$ km s$^{-1}$ velocity
dispersion. The one alteration we make to all of these templates is to
remove Ly$\alpha$, because it is generally much larger than observed
in real galaxies.

Let us refer to these 485 basis templates as $M_{j}(\lambda)$,
expressed in units of ergs s$^{-1}$ \AA$^{-1}$.

We seek to reduce this full basis space to a subspace in which
galaxies actually exist. In particular, we seek five templates
$F_{i}(\lambda)$ built from nonnegative combinations of the
original basis set of $N$ templates:
\begin{equation}
F_{i}(\lambda) = \sum_j b_{ij} M_{j}(\lambda),
\end{equation}
in units of ergs s$^{-1}$ \AA$^{-1}$. In principle, we could seek any
number of templates; from our experiments, we have found that five
turns out to be a number large enough to explain the data we use
here. For each galaxy $k$ we want our model
${\hat{F}_\lambda}(\lambda)$ for their spectrum to be a nonegative sum
of these five templates:
\begin{equation}
{\hat{F}}_{k}(\lambda) = \sum_i a_{ki} F_{i}(\lambda).
\end{equation}

@@

We begin with the simplest case, which is fitting a {\it single}
template to the photometric data of the LRG sample
(\citealt{eisenstein01a}). For many of these galaxies, which extend to
$r<19.5$ and are intrinsically red, the $u$-band flux is extremely
poorly measured, so we ignore the $u$-band for all LRGs.

Figure \ref{spec_lrg} shows the spectrum of the best-fit LRG.  This
template is constrained by data between about 2000 \AA\ and 10000 \AA;
outside that range it is an extrapolation.

@@

These templates explain the photometric data rather well. For example,
consider Figure \ref{fullfits}, showing the color residuals of the
observations with respect to the best fits, when fitting to galaxies
with GALEX, SDSS, and 2MASS data. In this context we define the color
residuals as (for example):
\begin{equation}
\Delta [u-g] = [u_{\mathrm{obs}}- g_{\mathrm{obs}}] -
[u_{\mathrm{model}}- g_{\mathrm{model}}] .
\end{equation}
In the code described in Section \ref{summary}, we used the routine {\tt
gst\_kcorrect} to perform these fits.  In Figure \ref{fullfits} the
thick dashed lines show the estimated 1$\sigma$ uncertainties in the
colors from the photometric catalogs. Relative to the uncertainties,
there are no significant biases or redshift trends in these fits. Note
that the galaxies used in this test are distinct from the training set
on which we fit the five templates.

The templates do well on higher redshift data, as well. For example,
Figure \ref{goods} shows the color residuals for GOODS data. Here we
find that there are somewhat larger residuals with respect to the
uncertainties (again shown as the thick dashed lines). We can compare
this result to that in Figure \ref{goods_special}, which we get by
fitting a new set of templates to {\it only} the GOODS data (this
alternate set of five templates is also distributed in the manner that
Appendix \ref{format} describes). From Figure \ref{goods_special}, we
conclude that {\it some} of the errors are irreducible, and require
either a larger set of templates for high redshift galaxies (a
possibility we find a priori unlikely), or that there are simply
errors in the catalogs or the input filter curves. For example, in the
$z$ and $V$ bands we find significant trends between redshifts $z=0.5$
and $1.5$ (echoed in Figure \ref{goods}). More significantly, in the
$H$ band there is at least a 10\% offset in the magnitudes across all
redshifts, which is almost certainly a catalog or filter curve
error. Other errors (such as a large scatter in the residuals in the
$B$ band and some redshift dependence in the $J$ band) are clearly due
to the fact that the templates are not primarily designed for GOODS
data. All in all, we recommend using the special templates for
$K$-corrections within the GOODS data set.

@@

In addition to $K$-corrections, these template fits also provide
physical interpretations of the galaxies, since they correspond to
actual stellar population synthesis models. Indeed, the code outputs
three parameters relevant to the star-formation history: the current
mass of stars in the galaxy; the stellar mass-to-light ratio of the
galaxy; and the fraction of the total amount of star-formation that
has occurred recently.

@@

Finally, we can also calculate measures of the recent star-formation
rate from these fits. Since our fits are to broad-band photometry, and
not based on emission line measurements, we cannot expect to have
detailed information about the very recent star-formation
rate. However, we can try to measure what fraction of the total
star-formation has occurred in, say, the past 1 Gyr:
\begin{equation}
b_{G} = \frac{\int_0^{1\mathrm{~Gyr}} dt\, \mathrm{SFR}(t)}
{\int_0^{a_{\mathrm{Univ}}} dt\, \mathrm{SFR}(t)}
\end{equation}
where the definition here is meant to be similar to the birthrate
parameter $b$ in \citet{kennicutt94a} and references therein.  Note
that here $b_G$ is defined in terms of the total integrated
star-formation history (unlike the stellar masses referred to above)
as is standard for the $b$ parameter. Figure \ref{umr_bg} shows the
relationship between the rest-frame $u-r$ color for SDSS galaxies and
$b_G$. For galaxies with $u-r < 2.5$ (redder than which galaxies are
mostly reddened, edge-on disks) these two parameters follow a simple
relationship:
\begin{equation}
\log_{10} b_g = -0.55 (u-r). 
\end{equation}
Note, however, the large scatter (about 0.3 dex at 1$\sigma$).

We do not want to overstate the validity of these physical
interpretations. They are sufficiently good physical interpretations
to explain broad band data, but are not uniquely so. Certainly more
detailed spectroscopic analysis can provide better constraints on
star-formation histories and stellar masses.

@@

\begin{thebibliography}{39}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

@@

@@

@@

@@

@@

\clearpage 
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{f17.eps}
\caption{\label{mass_to_garching} Our galaxy stellar mass estimates
	$M_\ast$ compared to those of \citet{kauffmann03a} $M_{s,\ast}$, as
	a function of stellar mass (top panel) and of color (bottom
	panel). The greyscale is the conditional distribution
	$M_{s,\ast}/M_\ast$ on each quantity . The lines are the 10\%, 25\%,
	50\%, 75\% and 90\% quantiles.}
\end{figure}

@@

\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{f19.eps}
\caption{\label{umr_bg} Fraction of the total star-formation that has
	occurred in the previous billion years $b_G$, as a function of restframe
	$u-r$ color, for SDSS galaxies. The greyscale is the conditional
	distribution of $b_G$ on $u-r$. The lines are the 10\%, 25\%, 50\%,
	75\% and 90\% quantiles. For $u-r < 2.5$, the median relationship
	follows the simple form listed in the figure. Galaxies with $u-r >
	2.5$ are often highly reddened star-forming galaxies.}
\end{figure}

>>>>

We begin with a basis set of 485 spectral templates. Of these, 450 are
a set of instantaneous bursts from \citet{bruzual03a}, using the
\citet{chabrier03a} stellar initial mass function and the Padova1994
isochrones. We use all 6 metallicities (mass fractions of elements
heavier than He of $Z= 0.0001$, $0.0004$, $0.004$, $0.008$, $0.02$ and
$0.05$).  For each metallicity we select 25 ages (between 1 Myr and
13.75 Gyr, spaced almost logarithmically in age).  We smooth each
template to $300$ km s$^{-1}$ velocity dispersion (we will smooth the
SDSS spectra in the training set to the same resolution).

There are limitations to all stellar population synthesis libraries,
and our results are therefore restricted to be at best as good as our
input models. There are some well-known areas of uncertainty in the
stellar populations models in the UV and IR. In the UV, old
populations sometimes contain extreme horizontal branch stars that
have lost their envelopes and are thus extremely hot, that dominate
the flux at $\lambda \sim 1500$ \AA. There is certainly no physical
model that predicts when this happens, and no consensus about how
these stars should be accounted for in isochrone synthesis (that is,
what fraction of stars undergo this phase as a function of mass and
metallicity). While the models of \cite{bruzual03a} do contain
post-AGB and planetary nebula phases that provide FUV flux in old
populations, they do not contain extreme horizontal branch stars or
other candidates that might provide more of the ``UV-excess'' seen in
elliptical galaxies (\citealt{greggio90a, brown00a}).  In the near-IR,
the spectra of thermally-pulsating asymptotic giant branch (TP-AGB)
stars may dominate the flux in certain intermediate age populations
(2--3 Gyr). With respect to these phases, \citet{bruzual03a} use
models that are calibrated to observations of TP-AGB stars
nearby. \citet{maraston05a} show that a different choice of TB-AGB
models can produce significantly higher flux in the NIR at
intermediate ages. Thus, the choices for AGB and post-AGB evolution we
make here by using the models \cite{bruzual03a} are not definitive,
but we do not address these controversies here and they remain an
uncertainty in the physical interpretation of galaxy fluxes.  We do
note that for low redshift galaxies our results from BC03 do explain
most galaxies' colors (though see below for our discussion of the high
redshift GOODS sample).

For each age and metallicity we make 3 choices of dust model: (1) no
dust extinction; (2) $\tau_V = 3$ dust with Milky Way type extinction
(as input into the models of \citealt{witt00a} with a ``homogeneous''
distribution and ``shell'' geometry); (3) $\tau_V = 3$ dust with SMC
type extinction (from the same models).

The remaining 35 templates are from MAPPINGS-III (\citealt{kewley01a})
models of emission from ionized gas. We choose the predictions for an
8 Myr old continuous star-formation history with 5 possible
metalliciticites ($Z=0.001$, 0.004, 0.008, 0.02, and 0.04) and 7
possible ionization parameters ($q= 5\times 10^6$, $10^7$, $2\times
10^7$, $4\times 10^7$, $8 \times 10^7$, $1.5\times 10^8$, and $3\times
10^8$ cm s$^{-1}$). We take the spectra (given as a set of delta
functions) and smooth them to $300$ km s$^{-1}$ velocity
dispersion. The one alteration we make to all of these templates is to
remove Ly$\alpha$, because it is generally much larger than observed
in real galaxies. We warn readers that galaxies exhibit many
individual variations in equivalent widths of emission lines --- our
final set of five templates will not have any hope of following these
variations in detail.

Let us refer to these 485 basis templates as $M_{j}(\lambda)$,
expressed in units of ergs s$^{-1}$ \AA$^{-1}$.  We seek to reduce
this full basis space to a subspace in which galaxies actually
exist. In particular, we seek five templates $F_{i}(\lambda)$ built
from nonnegative combinations of the original basis set of $N$
templates:
\begin{equation}
F_{i}(\lambda) = \sum_j b_{ij} M_{j}(\lambda),
\end{equation}
in units of ergs s$^{-1}$ \AA$^{-1}$. In principle, we could seek any
number of templates; from our experiments, we have found that five
turns out to be a number large enough to explain the data we use
here. For each galaxy $k$ we want our model
${\hat{F}_\lambda}(\lambda)$ for their spectrum to be a nonegative sum
of these five templates:
\begin{equation}
{\hat{F}}_{k}(\lambda) = \sum_i a_{ki} F_{i}(\lambda).
\end{equation}

@@

We begin with the simplest case, which is fitting a {\it single}
template to just the SDSS photometric data of the LRG sample
(\citealt{eisenstein01a}). For many of these galaxies, which extend to
$r<19.5$ and are intrinsically red, the $u$-band flux is extremely
poorly measured, so we ignore the $u$-band for all LRGs and only fit
to $g$, $r$, $i$ and $z$ fluxes using the SDSS ``model'' fluxes as
described in \cite{stoughton02a} and elsewhere.
In this case, we fit for the $a_{i0}$ and $b_{0j}$ coefficients 
in Equation \ref{chi2}: we want just a single template, expressed 
by the $b_{0j}$ coefficients, and a single scaling of its 
flux for each galaxy, expressed by $a_{i0}$.

Figure \ref{spec_lrg} shows the spectrum of the best-fit LRG.  This
template is constrained by data between about 3000 \AA\ and 10000 \AA.
Outside that range it is an extrapolation --- that is, it is simply
the best fit model prediction, but there is no data to verify the
model there.

@@

These templates explain the photometric data rather well. For example,
consider Figure \ref{fullfits}, showing the color residuals of the
observations with respect to the best fits, when fitting to galaxies
with GALEX, SDSS, and 2MASS data (using analogous sets of data to 
those described in Section \ref{data}, but in detail a different
set of galaxies). In this context we define the color
residuals as (for example):
\begin{equation}
\Delta [u-g] = [u_{\mathrm{obs}}- g_{\mathrm{obs}}] -
[u_{\mathrm{model}}- g_{\mathrm{model}}] .
\end{equation}
In the code described in Section \ref{summary}, we used the routine {\tt
gst\_kcorrect} to perform these fits.  In Figure \ref{fullfits} the
thick dashed lines show the estimated 1$\sigma$ uncertainties in the
colors from the photometric catalogs. Relative to the uncertainties,
there are no significant biases or redshift trends in these fits. Note
that the galaxies used in this test are distinct from the training set
on which we fit the five templates.

The templates do well on higher redshift data, as well. For example,
Figure \ref{goods} shows the color residuals for GOODS data. Here we
find that there are somewhat larger residuals with respect to the
uncertainties (again shown as the thick dashed lines). We can compare
this result to that in Figure \ref{goods_special}, which we get by
fitting a new set of templates to {\it only} the GOODS data (this
alternate set of five templates is also distributed in the manner that
Appendix \ref{format} describes). From Figure \ref{goods_special}, we
conclude that {\it some} of the errors are irreducible, and require
either a different set of templates for high redshift galaxies or that
there are simply errors in the catalogs or the input filter
curves. For example, in the $z$ and $V$ bands we find significant
trends between redshifts $z=0.5$ and $1.5$ (echoed in Figure
\ref{goods}). More significantly, in the $H$ band there is at least a
10\% offset in the magnitudes across all redshifts, which is almost
certainly a catalog or filter curve error. One possibility for these
problems in the red are that intermediate age populations have flux in
the near-IR from TP-AGB that our model omits; for example,
\citet{maraston05a} has shown that such stars can significantly change
the near-IR colors. We defer a fuller investigation of these
discrepancies.  Other errors (such as a large scatter in the residuals
in the $B$ band and some redshift dependence in the $J$ band) are
clearly due to the fact that the templates are not primarily designed
for GOODS data. All in all, we recommend using the special templates
for $K$-corrections within the GOODS data set.

@@

In addition to $K$-corrections, these template fits also provide
physical interpretations of the galaxies, since they correspond to
actual stellar population synthesis models. Indeed, the code outputs
three parameters relevant to the star-formation history: the current
mass of stars in the galaxy; the stellar mass-to-light ratio of the
galaxy; and the fraction of the total amount of star-formation that
has occurred recently. We do not believe that fitting just to
broad-band colors can ever give extremely accurate values for these
quantities, but to evaluate what {\tt kcorrect} does in this regard we
here examine the consistency between our results and those of other
groups for the same galaxies.

@@

Finally, we can also calculate measures of the recent star-formation
rate from these fits. Since our fits are to broad-band photometry, and
not based on emission line measurements, we cannot expect to have
detailed information about the very recent star-formation
rate. However, we can try to measure what fraction of the total
star-formation has occurred in, say, the past 300 Myr:
\begin{equation}
b_{300} = \frac{\int_0^{300\mathrm{~Myr}} dt\, \mathrm{SFR}(t)}
{\int_0^{a_{\mathrm{Univ}}} dt\, \mathrm{SFR}(t)}
\end{equation}
where the definition here is meant to be similar to the birthrate
parameter $b$ in \citet{kennicutt94a} and references therein.  Note
that here $b_{300}$ is defined in terms of the total integrated
star-formation history (unlike the stellar masses referred to above)
as is standard for the $b$ parameter. Figure \ref{umr_bg} shows the
relationship between the rest-frame $u-r$ color for SDSS galaxies and
$b_{300}$. For our results, the median relationship is close to:
\begin{equation}
\log_{10} b_{300} = 0.2 - 1.3 (u-r). 
\end{equation}
Note, however, the large scatter (about 0.3 dex at 1$\sigma$).  The
overlaid points show the results of using the formula of \citet{hopkins03a}
to infer the current star-formation rate from the $u$-band,
integrating over 300 Myr, scaling by the integrated stellar mass from
the {\tt kcorrect} determination, and then correcting for the internal
$u$-band dust extinction inferred by the {\tt kcorrect} fits (which
turns out to be factors ranging from 3--10, similar to the factors
used by \citealt{hopkins03a}).  Our results roughly scale as those of
\citet{hopkins03a} do, but with lower star-formation rates by a factor
of five or so. Since \citet{hopkins03a} calibrate their star-formation
rates to a number of other indicators, we conclude that either {\tt
kcorrect} probably underestimates the recent star-formation rate.

We do not want to overstate the validity of these physical
interpretations. They are sufficiently good physical interpretations
to explain broad band data, but are not uniquely so. Certainly more
detailed spectroscopic analysis can provide better constraints on
star-formation histories and stellar masses.  However, the results
from {\tt kcorrect} are roughly in agreement with what other authors
find with comparable data.

@@

\begin{thebibliography}{43}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

@@

\bibitem[{{Brown} {et~al.}(2000){Brown}, {Bowers}, {Kimble}, {Sweigart}, \&
  {Ferguson}}]{brown00a}
{Brown}, T.~M., {Bowers}, C.~W., {Kimble}, R.~A., {Sweigart}, A.~V., \&
  {Ferguson}, H.~C. 2000, \apj, 532, 308

@@

\bibitem[{{Greggio} \& {Renzini}(1990)}]{greggio90a}
{Greggio}, L. \& {Renzini}, A. 1990, \apj, 364, 35

@@

\bibitem[{{Hopkins} {et~al.}(2003){Hopkins}, {Miller}, {Nichol}, {Connolly},
  {Bernardi}, {G{\'o}mez}, {Goto}, {Tremonti}, {Brinkmann}, {Ivezi{\'c}}, \&
  {Lamb}}]{hopkins03a}
{Hopkins}, A.~M., {Miller}, C.~J., {Nichol}, R.~C., {Connolly}, A.~J.,
  {Bernardi}, M., {G{\'o}mez}, P.~L., {Goto}, T., {Tremonti}, C.~A.,
  {Brinkmann}, J., {Ivezi{\'c}}, {\v Z}., \& {Lamb}, D.~Q. 2003, \apj, 599, 971

@@

\bibitem[{{Maraston}(2005)}]{maraston05a}
{Maraston}, C. 2005, \mnras, 362, 799

@@

\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{f17.eps}
\caption{\label{mass_to_garching} Our galaxy stellar mass estimates
	$M_\ast$ compared to those of \citet{kauffmann03a} $M_{s,\ast}$, as
	a function of stellar mass (top panel) and of color (bottom
	panel). The greyscale is the conditional distribution
	$M_{s,\ast}/M_\ast$ on each quantity . The lines are the 10\%, 25\%,
	50\%, 75\% and 90\% quantiles.}
\end{figure}

@@

\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{f19.eps}
\caption{\label{umr_bg} Fraction of the total star-formation that has
	occurred in the previous 300 Myr $b_{300}$, as a function of
	restframe $u-r$ color, for SDSS galaxies. The greyscale is the
	conditional distribution of $b_{300}$ on $u-r$. The lines are the 10\%,
	25\%, 50\%, 75\% and 90\% quantiles.  For comparison we show the
	results of using the scaling of \citet{hopkins03a}.}
\end{figure}
